
title: "Why People Buy? A Statistic Analysis for Online Shopping Session"
author: |
  Kimiya Mavaddat\
  GH1043877
output:
  pdf_document:
    toc: true
    number_sections: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.width = 6, fig.height = 4
)
library(tidyverse)
library(car)      # leveneTest, vif
library(pROC)     # ROC, AUC
library(broom)    # for clean model result
library(scales)   # percent
theme_set(theme_minimal(base_size = 11))
```

# Business Problem

An online shop want to make bigger their **conversion rate** of the
website, I mean, the number of session that people buy something at the
end. The marketing team and product team need to know which behavior in a
session and which type of visitor is connect with buying, so the money for
personalize the website, re-targeting, and campaign in special season can
go for the factor that really make change, and not just spread same
amount everywhere in website.

For this analysis I am using the **Online Shoppers Purchasing Intention
Dataset** (Sakar et al., 2019), you can find it on Kaggle:
<https://www.kaggle.com/datasets/imakash3011/online-shoppers-purchasing-intention-dataset>.
It have 12,330 browsing session on a shopping website, in one year time,
each one have information about page-visit number and duration, engagement
number (`BounceRates`, `ExitRates`, `PageValues`), visitor and session
information (`Month`, `VisitorType`, `Weekend`, `TrafficType`, ...), and a
yes/no result `Revenue` that tell if the session end with purchase or not.

## Business questions

1. The session that convert, do they show **more engagement on the page**
   (`PageValues`) compare with the session that not convert?
2. Is buying connect to what type of visitor it is (new one, returning
   one, or other)?
3. If people browsing in the weekend, does it have relation with how much
   they buy?
4. Which factor, if we look them together and adjust for each other, can
   explain best the chance of purchase, and how good we can predict this
   chance?

# Exploratory Data Analysis

```{r load-data}
df_raw <- read.csv("data/online_shoppers_intention.csv", stringsAsFactors = FALSE)
dim(df_raw)
str(df_raw)
```

```{r missingness}
colSums(is.na(df_raw))
```

I check and there is no missing value in any column. So dataset is
complete already, and I don't need to do imputation for this one.

## Data cleaning

```{r duplicates}
n_dupe <- sum(duplicated(df_raw))
n_dupe
```

`r n_dupe` rows are **exact same as another row**, in all of the 18
columns, even in the number column like `BounceRates`/`ExitRates` that
have many decimal number after the point. I think it is not really
possible two different session have exact same number in every column
until that many decimal, so I treat this like the row get logged two time
or scraped two time by mistake, not like real independent session, and I
remove them. If I don't remove, the variance will look smaller than it
really is, and it will break the assumption of independent that the
hypothesis test need later.

```{r clean-data}
# June is full word here, not "Jun", so no month.abb
month_order <- c("Jan", "Feb", "Mar", "Apr", "May", "June",
                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

df <- df_raw %>%
  distinct() %>%
  mutate(
    Month        = factor(Month, levels = month_order[month_order %in% unique(Month)]),
    VisitorType  = factor(VisitorType),
    Weekend      = factor(Weekend, levels = c(FALSE, TRUE), labels = c("Weekday", "Weekend")),
    Revenue      = factor(Revenue, levels = c(FALSE, TRUE), labels = c("No purchase", "Purchase")),
    OperatingSystems = factor(OperatingSystems),
    Browser      = factor(Browser),
    Region       = factor(Region),
    TrafficType  = factor(TrafficType)
  )

dim(df)
stopifnot(sum(is.na(df$Month)) == 0)  # no NA from month change
```

**Outliers.** The duration and rate columns (`ProductRelated_Duration`,
`BounceRates`, `ExitRates`, `PageValues`) are very skew to the right side,
with lot of value on zero (session with only one page, or session with no
bounce/exit at all). I think this is just how people really shop, not a
mistake when someone type the data, so I don't remove any row for this. I
handle the skew instead with log transform, in the place it matter for the
modeling part (Section 5).

## Descriptive statistics

```{r desc-numeric}
numeric_vars <- c("Administrative", "Administrative_Duration", "Informational",
                   "Informational_Duration", "ProductRelated", "ProductRelated_Duration",
                   "BounceRates", "ExitRates", "PageValues", "SpecialDay")

df %>%
  select(all_of(numeric_vars)) %>%
  summarise(across(everything(),
                    list(mean = ~mean(.x), median = ~median(.x),
                         sd = ~sd(.x), IQR = ~IQR(.x)))) %>%
  pivot_longer(everything(), names_to = "stat", values_to = "value") %>%
  separate(stat, into = c("variable", "statistic"), sep = "_(?=[^_]+$)") %>%
  pivot_wider(names_from = statistic, values_from = value) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))
```

```{r desc-categorical}
table(df$Revenue) %>% prop.table() %>% round(3)
table(df$VisitorType)
table(df$Weekend)
```

Purchase is the **small group here**: only around 15% of session end with
buying. I keep this imbalance in my mind for later, when I look at the
proportion and also the model performance (Section 5.4).

## Purchase rate by segment

```{r eda-plots, fig.height=3.2}
p1 <- df %>%
  ggplot(aes(x = Revenue, y = PageValues)) +
  geom_boxplot(outlier.alpha = 0.15, fill = "#4C78A8") +
  coord_cartesian(ylim = c(0, 60)) +
  labs(title = "PageValues, by if purchase happen or not", x = NULL, y = "PageValues")

p2 <- df %>%
  count(VisitorType, Revenue) %>%
  group_by(VisitorType) %>%
  mutate(rate = n / sum(n)) %>%
  filter(Revenue == "Purchase") %>%
  ggplot(aes(x = VisitorType, y = rate)) +
  geom_col(fill = "#4C78A8") +
  scale_y_continuous(labels = percent) +
  labs(title = "Purchase rate for every visitor type", x = NULL, y = "Purchase rate")

p1
p2
```

```{r eda-plots-2, fig.height=3.2}
p3 <- df %>%
  count(Weekend, Revenue) %>%
  group_by(Weekend) %>%
  mutate(rate = n / sum(n)) %>%
  filter(Revenue == "Purchase") %>%
  ggplot(aes(x = Weekend, y = rate)) +
  geom_col(fill = "#4C78A8") +
  scale_y_continuous(labels = percent) +
  labs(title = "Purchase rate, weekday against weekend", x = NULL, y = "Purchase rate")

p4 <- df %>%
  count(Month, Revenue) %>%
  group_by(Month) %>%
  mutate(rate = n / sum(n)) %>%
  filter(Revenue == "Purchase") %>%
  ggplot(aes(x = Month, y = rate)) +
  geom_col(fill = "#4C78A8") +
  scale_y_continuous(labels = percent) +
  labs(title = "Purchase rate in every month", x = NULL, y = "Purchase rate")

p3
p4
```

```{r eda-correlation, fig.height=4.5}
cor_mat <- df %>% select(all_of(numeric_vars)) %>% cor()
cor_df <- as.data.frame(as.table(cor_mat))

ggplot(cor_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient2(low = "#B00020", mid = "white", high = "#4C78A8",
                        midpoint = 0, limits = c(-1, 1), name = "corr") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation between number columns", x = NULL, y = NULL)
```

`BounceRates` and `ExitRates` have very strong correlation together (make
sense, both of them talking about session where person leave fast). I come
back to this when I choose predictor for the regression model in Section
5, because I want to avoid the multicollinearity problem.

# Hypotheses and Test Selection

```{r stratified-split}
set.seed(42)
train_idx <- unlist(lapply(split(seq_len(nrow(df)), df$Revenue),
                            function(idx) sample(idx, size = floor(0.8 * length(idx)))))
train <- df[train_idx, ]
test  <- df[-train_idx, ]

c(train = nrow(train), test = nrow(test))
prop.table(table(train$Revenue))
prop.table(table(test$Revenue))
```

I make a **stratified 80/20 split** (with seed = 42, fix number so result
is same every time) so the purchase rate stay same amount in both group.
The hypothesis test number 1 to 3 below, I run on the whole clean dataset
(no model fitting needed for them, so no need train/test split there); the
`train` group is for fit the logistic regression in Section 5, and `test`
group I only use for check how good it is.

## H1: Engagement (`PageValues`) and Purchase

$H_0$: distribution of `PageValues` is same for session that purchase and
session that not purchase. $H_1$: it is not same.

```{r h1-assumptions, fig.height=3}
qqnorm(df$PageValues[df$Revenue == "Purchase"], main = "QQ plot, PageValues (Purchase)")
qqline(df$PageValues[df$Revenue == "Purchase"])

set.seed(42)
shapiro.test(sample(df$PageValues, 5000))
leveneTest(PageValues ~ Revenue, data = df)
```

The QQ plot show big jump away from the straight line, and Shapiro-Wilk
test on 5,000-row sample (biggest sample size this test can take) say the
data is not normal ($p < .001$). Levene's test also say the variance is
not equal. Because both of the thing that parametric t-test need are
broken, I use the **Mann-Whitney U test** instead (it use rank, and it
don't need normal distribution) as my main test, and I still show Welch's
t-test next to it, just to double check.

```{r h1-test}
wilcox.test(PageValues ~ Revenue, data = df)
t.test(PageValues ~ Revenue, data = df)
```

**What I think this mean:** both test say reject $H_0$ at $\alpha = .05$
($p < 2.2\times10^{-16}$). The session that end in purchase have higher
`PageValues` number than the session that not purchase, and this is
happen systematically, not by luck. This match with the fact that
`PageValues` (a number that Google Analytics give to a page, based on how
often that page come right before someone buy) is a strong signal for
buying intent.

## H2: Visitor Type and Purchase

$H_0$: `Revenue` and `VisitorType` don't have relation. $H_1$: they have
relation.

```{r h2-test}
tab_visitor <- table(df$VisitorType, df$Revenue)
tab_visitor
chi_visitor <- chisq.test(tab_visitor)
chi_visitor
chi_visitor$expected  # rule: all numbers > 5
```

Every expected number in the cell is bigger than 5, so the chi-square
approximation should work fine here. The test say reject $H_0$
($p < 2.2\times10^{-16}$), so purchase behavior have relation with visitor
type. If I look at the descriptive number (Section 2.4), I can see
**new visitor buy with higher rate than returning visitor**, and this look
strange at first, but maybe it make sense because returning visitor is
just browsing or checking thing again, without real need to buy right now.

## H3: Weekend and Purchase

$H_0$: `Revenue` and `Weekend` don't have relation. $H_1$: they have
relation.

```{r h3-test}
tab_weekend <- table(df$Weekend, df$Revenue)
tab_weekend
chisq.test(tab_weekend)
```

The test say reject $H_0$ ($p \approx$ `r signif(chisq.test(tab_weekend)$p.value, 3)`):
session in weekend buy with a little bit higher rate than session in
weekday, but this effect is much more small if I compare with `PageValues`
or `VisitorType` (I talk more about effect size in Section 6).

# Logistic Regression: Joint Drivers of Purchase

Hypothesis 1 to 3 only test one factor at one time. To see which factor
still matter **after we adjust for the other one**, and also to get model
that can give real *probability* number for purchase, not only tell if
there is relation, I fit a multiple logistic regression on the `train`
group.

## Predictor selection

`Region`, `Browser`, `OperatingSystems`, and `TrafficType` are just
anonymous number code, with many level and no clear business meaning, so
if I put them in the model it just add many coefficient with no useful
story behind, and because of this I take them out. `BounceRates` I also
take out and keep `ExitRates` only, because of the strong correlation I
saw already in Section 2.5 (if I keep both, it will make variance bigger
for no reason, without giving new information).

`PageValues` and `ProductRelated_Duration` are skew to right side (Section
2), so I use `log1p()` on both of them. I check below, in empirical way,
by comparing the model with transform and without transform.

```{r logit-transform-check}
m_raw <- glm(Revenue ~ PageValues + ProductRelated_Duration, data = train, family = binomial)
m_log <- glm(Revenue ~ log1p(PageValues) + log1p(ProductRelated_Duration), data = train, family = binomial)
AIC(m_raw, m_log)
```

The log-transform version have lower AIC number, so this confirm the
transform is making the fit better, and I use it in the full model.

```{r logit-fit}
month_mode <- names(sort(table(train$Month), decreasing = TRUE))[1]
train$Month <- relevel(train$Month, ref = month_mode)
test$Month  <- factor(test$Month, levels = levels(train$Month))

logit_model <- glm(
  Revenue ~ log1p(PageValues) + VisitorType + Weekend + Month +
            log1p(ProductRelated_Duration) + ExitRates + SpecialDay,
  data = train, family = binomial
)
summary(logit_model)
```

## Assumption checks

```{r logit-vif}
vif(logit_model)
```

All the GVIF$^{1/(2\cdot df)}$ number are well under the common line of 2,
so multicollinearity is not really a problem for this group of predictor.

```{r logit-linearity, fig.height=3.2}
train %>%
  mutate(logit = predict(logit_model, type = "link")) %>%
  ggplot(aes(x = log1p(PageValues), y = logit)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "loess", se = FALSE, color = "#B00020") +
  labs(title = "Check for linearity: logit and log1p(PageValues)")
```

The loess line stay close to straight line for most of the range, so this
support the linear-in-the-logit idea for this predictor (there is some
curve at the very right side, where data point are not many, but this is
expected and not big deal).

## Coefficients as odds ratios

```{r logit-odds}
tidy(logit_model, exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))
```

**What the main effect mean (odds ratio, 95% CI):**

- Every time `log1p(PageValues)` go up by 1 unit, the odds of purchase get
  multiply by the exponent coefficient shown above. This is by far the
  biggest effect in whole model, and it confirm H1 for me.
- `VisitorType = New_Visitor` have higher odds for purchase, compare with
  the `Returning_Visitor` baseline, and this match with H2.
- The `Weekend` effect, after adjust (OR `r round(tidy(logit_model, exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "WeekendWeekend") %>% pull(estimate), 2)`,
  95% CI `r round(tidy(logit_model, exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "WeekendWeekend") %>% pull(conf.low), 2)`-`r round(tidy(logit_model, exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "WeekendWeekend") %>% pull(conf.high), 2)`,
  $p \approx$ `r round(tidy(logit_model, exponentiate = TRUE, conf.int = TRUE) %>% filter(term == "WeekendWeekend") %>% pull(p.value), 2)`)
  is **not** statistically significant, and this is different from what
  H3 bivariate chi-square test show us before. I think the raw relation
  between weekend and purchase, it is actually explain by other factor
  that already in the model, like which month or which visitor type is
  more active in weekend, and not an independent effect from weekend by
  itself. I think this is good example why adjusted model is important,
  and not only doing one-factor-at-a-time test, for this business
  question: it will be a mistake if we spend budget only for target
  "weekend shopper" as one segment, just because of H3 alone.
- `ExitRates` have strong negative effect: session where visitor tend to
  leave from a page, they have much lower odds to convert.

## the model performance and confidence :D

```{r logit-performance}
test$pred_prob <- predict(logit_model, newdata = test, type = "response")

roc_obj <- roc(test$Revenue, test$pred_prob, levels = c("No purchase", "Purchase"))
auc(roc_obj)

best_thresh <- coords(roc_obj, "best", ret = "threshold", best.method = "youden")
best_thresh

pred_class_05  <- factor(ifelse(test$pred_prob > 0.5, "Purchase", "No purchase"),
                          levels = levels(test$Revenue))
pred_class_opt <- factor(ifelse(test$pred_prob > as.numeric(best_thresh), "Purchase", "No purchase"),
                          levels = levels(test$Revenue))

cat("Confusion matrix, at 0.5 cut point:\n")
table(predicted = pred_class_05, actual = test$Revenue)

cat("\nConfusion matrix, at the best Youden cut point:\n")
table(predicted = pred_class_opt, actual = test$Revenue)

mcfadden_r2 <- 1 - (logit_model$deviance / logit_model$null.deviance)
mcfadden_r2
```

```{r roc-plot, fig.height=3.5}
plot(roc_obj, main = paste0("ROC curve (AUC = ", round(auc(roc_obj), 3), ")"))
```

The model get an **AUC of `r round(auc(roc_obj), 3)`**, this mean if I pick
one random purchase session and one random non-purchase session, the model
rank the purchase one higher, that often. McFadden pseudo-$R^2$ is
`r round(mcfadden_r2, 3)`, which is not a big number but also not small,
and this is normal for this kind of behavior data because people behavior
have a lot of randomness in it. At the 0.5 cut point, the small group
(`Purchase`) get under-predict because of the class imbalance we talk
about (Section 2.3). The Youden cut point trade a little precision for
much better recall on the purchase, and I think this is usually the more
important error to avoid for business (if we miss a real buyer, it cost
more than showing one extra ad to someone who not going to buy anyway).

